---
title: "``varSelection`` function"
author: "Simone Del Sarto"
date: "21/7/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This function allows for the item selection based on two steps:

1. preliminary selection (non model-based)
2. IRT selection (model-based)

## Preparation
```{r }
source("varSelection.R")
load("dati_UNHCR.RData")
```
where "dati_UNHCR.RData" is an object containing the data for which the item selection has to be applied.

Suppose the ``household`` object contained the dataset at issue (UNHCR Mauritania household dataset)

```{r }
nrow(household); ncol(household)
```

First, select the food sections (from 4 to 10): columns from 20 to 109

```{r}
hh <- household[, 20:109] #90 variables
```

This will be the first argument of function ``varSelection``.

## Arguments
```{r echo=FALSE}
args(varSelection)
```

* ``data``: dataset (dichotomous variable must be 0/1 coded)
* ``miss_thr``: threshold for missing values (default 2/3)
* ``gamma_thr``: threshold for item discrimination (default 0.7)
* ``beta_thr``: threshold for beta parameters (default [-3,3])
* ``m``: max no. of categories for discrete variables, otherwise it is considered as continuous
* ``crit_thr``: threshold for "critical items"; observed response rate concentrated in a certain category (default 0.95)
* ``Theta_range``: latent trait range (standard normal distributed) for evaluating the item/test information (default [-15,15])
* ``qsel``: quantile for the selection based on item information proportion on the whole test information (default 0.25 = 1st quartile)
* ``...``: further (if needed) arguments for 'mirt' function within IRT selection

```{r }
out <- varSelection(hh, verbose=FALSE)
```

The output is a list of two objects:
```{r }
str(out$prel)
```

It contains the summary of all the variables in the initial dataset

* ``ID``: ID of the variables
* ``name``: name of the variables
* ``miss_prop``: observed proportion of missing values
* ``ncat``: number of categories (observed)
* ``min``: (observed) minimum values
* ``max``: (observed) maximum values
* ``mean``: observed average values
* ``cont``: whether the variables are considered continuous (``ncat`` $>$ ``m``), then discarded
* ``miss``: whether the missing values are beyond the threshold (``miss_thr``), then discarded

```{r }
str(out$final)
```

It contains the summary of the variables retained after the preliminary selection. In addition to the above summary, it has further columns:

* ``crit_item``: whether the items are considered as "critical" (observed response rate for a certain category $>$ ``crit_thr``)
* ``gamma``: estimated discrimination parameters
* ``beta1``, ``beta2``, ...: estimated difficulty (cut-off) parameters (as many as the number of response categories minus 1)
* ``prop_info``: proportion (%) of item information over the whole test information
* ``gamma_sel``: whether the item is kept or dropped according to the discrimination criterion ($\geq$ ``gamma_thr``)
* ``beta_sel``: whether the item is kept or dropped according to the beta criterion (the (median) beta parameter outside interval in ``beta_thr``)
* ``info_sel``: whether the item is kept or dropped according to the first criterion based on item information proportion ($\geq (1/J)\times 100$, where $J$ is the number of items)
* ``info_selQ``: whether the item is kept or dropped according to the second criterion based on item information proportion (item information proportion $\geq$ ``qsel``-quantile of the information proportion distribution)
* ``final_sel``: final selection -- whether the item is kept or dropped (at least two criteria out of four)
